---
title: "Integration of Hutchinson and Howick data"
author: "Virginia Howick"
date: "26/10/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---


```{r, message=FALSE}
setwd("/Users/virginiahowick/Documents/Tryps")
library(scater, quietly = TRUE)
library(scmap)
library(Seurat)
library(scater)
library(scran)
library(cowplot)
library(gridExtra)
library(viridis)
```

```{r, tidy=TRUE}
load("/Users/virginiahowick/Documents/Tryps/T1_3")
T1_3@meta.data$study <- rep("hutchinson", length(T1_3@meta.data$orig.ident))

#vi.seurat <- readRDS("/Users/virginiahowick/Documents/Tryps/data/vi_seurat_20201020.rds")

tca.qc.scran <- readRDS("howick_tryps_sce.rds")
tca.seurat <- as.Seurat(tca.qc.scran, counts = "counts", data = "logcounts")
Idents(tca.seurat) <- "seurat_clusters"
#tca.seurat <- subset(x = tca.seurat, idents = "SG")
tca.seurat <- FindVariableFeatures(tca.seurat, selection.method = "vst", nfeatures = 500)
tca.seurat@meta.data$study <- rep("howick", length(tca.seurat@meta.data$sample_id))
```


##Perform integration
We then identify anchors using the FindIntegrationAnchors function, which takes a list of Seurat objects as input, and use these anchors to integrate the two datasets together with IntegrateData.
```{r, tidy=TRUE}
p.anchors <- FindIntegrationAnchors(object.list = list(tca.seurat, T1_3), dims = 1:20)
p.combined <- IntegrateData(anchorset = p.anchors, dims = 1:20)
```

##Perform an integrated analysis
Now we can run a single integrated analysis on all cells!
```{r, tidy=TRUE}
DefaultAssay(p.combined) <- "integrated"

# Run the standard workflow for visualization and clustering
p.combined <- ScaleData(p.combined, verbose = FALSE)
p.combined <- RunPCA(p.combined, npcs = 30, verbose = FALSE)
# t-SNE and Clustering
#p.combined <- FindVariableFeatures(p.combined, selection.method = "vst", nfeatures = 500)
#hvg <- HVFInfo(object = p.combined)
#hvgfeat <- rownames(hvg)

p.combined <- RunUMAP(p.combined,  reduction = "pca", dims = 1:20, seed.use=222)
p.combined <- RunUMAP(p.combined,  reduction = "pca", dims = 1:20, umap.method = "uwot", n.neighbors = 5, min.dist = 2, spread = 3, seed.use=222)

p <- DimPlot(p.combined, reduction = "umap", group.by = "study")
p

pdat <- p$data

colors = c(howick = "#3399FF", hutchinson = "#FF9933")

ggplot(pdat, aes(UMAP_1, UMAP_2)) + geom_point(aes(colour = study), size=0.5) + theme_bw() + 
    scale_colour_manual(values = colors) +
  theme(axis.text = element_blank(), 
    axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
    axis.title = element_text(size = 10), legend.position = "none")

p <- DimPlot(p.combined, reduction = "umap", label=TRUE)

all2 <- p$data
colnames(all2) <- c("UMAP_1", "UMAP_2", "paper_ids")

pfcenters <- aggregate(all2[, 1:2], list(all2$paper_ids), mean)

colors = c(C1 = "red", C2 = "orange", C3 = "gold", C4 = "seagreen3", C5 = "turquoise4", 
    C6 = "#00CCFF", 'Attached Epimastigote' = "#6633FF", Gametes = "#CC00CC", Metacyclics = "purple", 'Midgut forms' = "cyan", 'Pre-metacyclic' = "coral")


ggplot(all2, aes(UMAP_1, UMAP_2)) + 
  geom_point(aes(colour = paper_ids), size=2) + 
  geom_text(aes(x=UMAP_1, y=UMAP_2, label=Group.1), data=pfcenters, hjust=0, vjust=0, size=2) +
  theme_bw() + 
    scale_colour_manual(values = colors, breaks = c("C1", "C2", "C3", "C4", 
        "C5", "C6", "Midgut forms", "Gametes", "Attached Epimastigote", "Pre-metacyclic", "Metacyclics")) + 
    theme(axis.text = element_blank(), 
    axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
    axis.title = element_text(size = 10), legend.position = "none")


p.combined@meta.data$paper_ids <- Idents(p.combined)

p.combined <- FindNeighbors(p.combined, reduction = "pca", dims = 1:20)
p.combined <- FindClusters(p.combined, resolution = 0.8)

DimPlot(p.combined, reduction = "umap")

p1 <- DimPlot(p.combined, reduction = "umap", group.by = "paper_ids", label = TRUE)
p2 <- DimPlot(p.combined, reduction = "umap", label = TRUE)
plot_grid(p1, p2)

all2 <- p2$data

pfcenters <- aggregate(all2[, 1:2], list(all2$ident), mean)


ggplot(all2, aes(UMAP_1, UMAP_2)) + 
  geom_point(aes(colour = ident), size=0.5) + 
  geom_text(aes(x=UMAP_1, y=UMAP_2, label=Group.1), data=pfcenters, hjust=0, vjust=0) +
  theme_bw() + 
    theme(axis.text = element_blank(), 
    axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
    axis.title = element_text(size = 10), legend.position = "none")

FeaturePlot(p.combined, "Tb927.10.12080")
FeaturePlot(T1_3, "Tb927.10.12080")



```

##Identify conserved cell type markers
To identify canonical cell type marker genes that are conserved across conditions, we provide the FindConservedMarkers function. This function performs differential gene expression testing for each dataset/group and combines the p-values using meta-analysis methods from the MetaDE R package.

```{r, tidy=TRUE}
DefaultAssay(p.combined) <- "RNA"
Idents(p.combined) <- "seurat_clusters"

gam.markers <- FindConservedMarkers(p.combined, ident.1 = 3, grouping.var = "study", verbose = FALSE, logfc.threshold = 0.25, only.pos=TRUE)
head(gam.markers)

anno <- read.csv("Tbb927_allanno.csv", header=TRUE)

gam.markers$gene_id <- rownames(gam.markers)

gam.markers$prod_desc <- anno[match(gam.markers$gene_id, anno[, 1]), 3]
gam.markers$gene_name <- anno[match(gam.markers$gene_id, anno[, 1]), 6]


#FeaturePlot(p.combined, features = c("ORTH-612", "ORTH-300", "ORTH-3513", "ORTH-1259"), min.cutoff = "q9", reduction = "umap")

pc.markers <- FindConservedMarkers(p.combined, ident.1 = 4, grouping.var = "study", verbose = FALSE, logfc.threshold = 0.25, only.pos=TRUE)
head(pc.markers)

pc.markers$gene_id <- rownames(pc.markers)

pc.markers$prod_desc <- anno[match(pc.markers$gene_id, anno[, 1]), 3]
pc.markers$gene_name <- anno[match(pc.markers$gene_id, anno[, 1]), 6]

FeaturePlot(p.combined, features = c("Tb927.3.4500", "Tb927.9.7470", "Tb927.4.4730", "Tb927.7.7090"), min.cutoff = "q9", reduction = "umap")



s <- FeaturePlot(p.combined, features = c("Tb927.3.4500"), min.cutoff = "q9", reduction = "umap", max.cutoff = 3,
    cols = c("grey", "red"))
sdat <- s$data
ps <- ggplot(sdat, aes(UMAP_1, UMAP_2)) + 
  geom_point(aes_string(colour = "Tb927.3.4500"), size=0.5) + 
  #labs(title="Tb927.3.4500") +
  scale_colour_gradient(low = "grey", high = "blue") +
  theme_bw() + 
  theme(axis.text = element_blank(), 
    axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
    axis.title = element_text(size = 10), legend.position = "none")

ps

f <- FeaturePlot(p.combined, features = c("Tb927.7.380"), min.cutoff = "q9", reduction = "umap", max.cutoff = 3,
    cols = c("grey", "red"))
fdat <- f$data
pf <- ggplot(fdat, aes(UMAP_1, UMAP_2)) + 
  geom_point(aes_string(colour = "Tb927.7.380"), size=0.5) + 
  #labs(title="Tb927.7.380", x=element_blank(), y=element_blank()) +
  scale_colour_gradient(low = "grey", high = "blue") +
  theme_bw() + 
  theme(axis.text = element_blank(), 
    axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
    axis.title = element_text(size = 10), legend.position = "none")

plot_grid(ps, pf)

h <- FeaturePlot(p.combined, features = c("Tb927.10.10770"), min.cutoff = "q9", reduction = "umap", max.cutoff = 3,
    cols = c("grey", "red"))
fdat <- h$data
pf <- ggplot(fdat, aes(UMAP_1, UMAP_2)) + 
  geom_point(aes_string(colour = "Tb927.10.10770"), size=0.5) + 
  #labs(title="Tb927.7.380", x=element_blank(), y=element_blank()) +
  scale_colour_gradient(low = "grey", high = "blue") +
  theme_bw() + 
  theme(axis.text = element_blank(), 
    axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
    axis.title = element_text(size = 10), legend.position = "none")
pf

#grid.arrange(pm, pf, po, ps, nrow=2, ncol=2)

subgam.markers <- gam.markers[gam.markers$max_pval < 0.001, ] 
subpc.markers <- pc.markers[pc.markers$max_pval < 0.001, ] 


saveRDS(p.combined, file="hutch_howick_combo.rds")
write.csv(subgam.markers, "hutch_howick_gam_markers.csv")
write.csv(subpc.markers, "hutch_howick_pc_markers.csv")


```

```{r}
p.sce <- as.SingleCellExperiment(p.combined)

subsubgam <- subgam.markers[1:10, ]
subsubgam$marker <- rep("gam", length(subsubgam$howick_p_val))
subsubpc <- subpc.markers[1:10, ]
subsubpc$marker <- rep("midgut", length(subsubpc$howick_p_val))

# all <- rbind(subsubpc, subsubgam)
# 
# p.sce.m <- p.sce[rownames(p.sce) %in% rownames(all), ]
# 
# lcpm_mat <- as.data.frame(assays(p.sce.m)[["logcounts"]])
# tmat <- t(lcpm_mat)
# tmat <- as.data.frame(tmat)
# 
# tmat$seurat_cluster <- as.factor(p.sce$ident)
# tmat$clust_paper <- paste(tmat$seurat_cluster, p.sce$study, sep="_")
# tmat <- as.data.frame(tmat)
# #lcpm_mat_sub <- lcpm_mat[!is.na(lcpm_mat$Cluster_k15), ]
# 
# clustmean <- aggregate(tmat[, 1:20], by = list(as.factor(as.character(tmat$clust_paper))), 
#     mean)
# rownames(clustmean) <- clustmean$Group.1
# 
# clustmean2 <- clustmean[, 2:21]
# 
# tcm2 <- as.data.frame(t(clustmean2))
# 
# 
# pheatmap(tcm2, show_colnames = TRUE, show_rownames = TRUE, cluster_cols = FALSE,
#     color = brewer.pal(9, "PuRd"))
```

```{r}
# m <- c(3,4)
# tmatsub <- tmat[tmat$seurat_cluster %in% m, ]
# 
# clustmean <- aggregate(tmatsub[, 1:20], by = list(as.factor(as.character(tmatsub$clust_paper))), 
#     mean)
# rownames(clustmean) <- clustmean$Group.1
# 
# clustmean2 <- clustmean[, 2:21]
# 
# tcm2 <- as.data.frame(t(clustmean2))
# 
# 
# pheatmap(tcm2, show_colnames = TRUE, show_rownames = TRUE, cluster_cols = FALSE,
#     color = brewer.pal(9, "PuRd"))
# ```
# 
# 
# ```{r}
# tcm3 <- tcm2
# 
# tcm3$gene_id <- rownames(tcm3)
# 
# tcm3$gene_name <- anno[match(tcm3$gene_id, anno[, 1]), 6]
# 
# rownames(tcm3) <- paste(tcm3$gene_id, tcm3$gene_name, sep="_")
# 
# tcm4 <- tcm3[, 1:4]

#pheatmap(tcm4, show_colnames = TRUE, show_rownames = TRUE, cluster_cols = FALSE,
 #   color = brewer.pal(9, "PuRd"))

#pheatmap(tcm4, show_colnames = TRUE, show_rownames = TRUE, cluster_cols = FALSE, cluster_rows = FALSE,
#color = brewer.pal(9, "PuRd"), gaps_row = 10, gaps_col = 2)

```

```{r}

# all$gene_anno <- paste(all$gene_id, all$gene_name, sep="_")
# tcm5 <- tcm3[match(all$gene_id, tcm3$gene_id), ]
# ann_c <- list(stage = cols)
# 
# 
# rownames(rd) <- rd$man_anno
# stage <- rd["stage"]


#pheatmap(tcm5, show_colnames = TRUE, show_rownames = TRUE, cluster_cols = FALSE, cluster_rows = FALSE,
 #   color = brewer.pal(9, "PuRd"), gaps_row = 10, gaps_col = c(2,4,6,8), annotation_row=stage, fontsize = 8, annotation_colors = ann_c)
```

